---
title: 'OTDM - Constrained Optimization - SVM'
author: "Julian Fransen, Danila Kokin"
date: "2024-11-24"
output:
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
editor_options:
  chunk_output_type: inline
---


1. A cover page with the name of the two (and exactly two!) members of the group

put a picture of UPC maybe? we have to knit to pdf i think

\begin{titlepage}
    \centering
    \vspace*{4cm} % Adjust vertical spacing
    \includegraphics[width=\textwidth]{./UPC.jpg} % Replace with your image path
    \vfill
    {\Huge \textbf{Your Report Title}} \\[1.5cm]
    {\Large Author: Your Name} \\[0.5cm]
    {\Large Date: \today}
    \vfill
\end{titlepage}


## 2. Description of Data Matrix A

### Title: Wine Data (Adapted for Unsupervised Learning)

### Sources:
- **Original Source**: UCI Machine Learning Repository
- **Original Dataset**: [Wine Dataset](https://archive.ics.uci.edu/ml/datasets/wine)
- **Creator**: Forina et al. (1991), Analytical Chemistry

### Relevant Information:
These data are the results of a chemical analysis of wines grown in the same region in Italy, derived from three different cultivars. The analysis determined the quantities of 13 constituents found in each of the three types of wines. The dataset is adapted for unsupervised learning by removing the labels that classify the types of wines.

- **Data Usage**: Frequently used for clustering, PCA, and classification studies.
- **Domain**: Chemistry, wine analysis, and machine learning.

### Number of Instances:
- 178 wine samples.

### Number of Attributes:
- 13 numeric continuous attributes.

### Attribute Information:
1. Alcohol
2. Malic acid
3. Ash
4. Alcalinity of ash
5. Magnesium
6. Total phenols
7. Flavanoids
8. Nonflavanoid phenols
9. Proanthocyanins
10. Color intensity
11. Hue
12. OD280/OD315 of diluted wines
13. Proline

### Missing Attribute Values:
- None.

### Preprocessing:
The dataset was scaled using Z-score normalization to standardize the features to zero mean and unit variance for better interpretability and clustering performance. After scaling, the Euclidean distance was calculated, and this matrix of distances was formatted to AMPL compatible .dat files.

### Summary Statistics:

```{r echo=FALSE}
summary_stats <- data.frame(
  Feature = c("Alcohol", "Malic acid", "Ash", "Alcalinity of ash", "Magnesium", 
              "Total phenols", "Flavanoids", "Nonflavanoid phenols", 
              "Proanthocyanins", "Color intensity", "Hue", 
              "OD280/OD315 of diluted wines", "Proline"),
  Min = c(11.03, 0.74, 1.36, 10.6, 70, 0.98, 0.34, 0.13, 0.41, 1.28, 0.48, 1.27, 278),
  Max = c(14.83, 5.80, 3.23, 30.0, 162, 3.88, 5.08, 0.66, 3.58, 13.0, 1.71, 4.00, 1680),
  Mean = c(13.00, 2.34, 2.36, 19.49, 99.74, 2.29, 2.03, 0.36, 1.59, 4.71, 0.96, 2.61, 746.89),
  `Standard Deviation` = c(0.81, 1.12, 0.27, 3.34, 14.28, 0.63, 0.99, 0.12, 0.57, 2.43, 0.23, 0.71, 314.91)
)

knitr::kable(summary_stats, format = "markdown", col.names = c("Feature", "Min", "Max", "Mean", "Standard Deviation"))
```

## 3. The AMPL `.mod` and `.dat` Files

### Sample of `wine.dat` File:

```plaintext
param m := 178;
param k := 3;
param d: 1 2 ... 177 178 :=
1   0.000000  3.497535 ...  6.078781  7.184421
2   3.497535  0.000000 ...  6.094927  7.367719
... 
178 7.184421  7.367719 ...  3.324276  0.000000;
```

### Contents of `clustering.mod` file:

```plaintext
/#Below we define the AMPL model for clustering
param m;
param k;
param d{1..m, 1..m};
var x{1..m, 1..m} binary;

minimize f: sum{i in 1..m, j in 1..m} d[i,j]*x[i,j];
subject to c1{i in 1..m}: sum{j in 1..m} x[i,j] = 1;
/# This constraint ensures every point belongs to 1 cluster
subject to c2: sum{j in 1..m} x[j,j] = k;
/# This constraint ensure that there are exactly k clusters
subject to c3{i in 1..m, j in 1..m}: x[j,j] >= x[i,j];
/# This constraint ensures that a point can only be part of a cluster that exists
```


## 4. The optimal solution obtained with AMPL.

```{bash, echo=FALSE}
cd /home/julian/uni_folder/OTDM/OTDM_Project_2/3/linprog/
/home/julian/uni_folder/OTDM/OTDM_p2/OTDM/ampl_linux-intel64/ampl << EOF 
reset;
reset data;

model clustering.mod;
data wine2.dat;
option solver cplex;
option cplex_options 'timing 1';
solve;

print "     Matrix x (formulation)"; 
print "";

print "     Number of points per cluster"; 
print "";

var total{i in 1..m};

for {i in 1..m} {
    for {j in 1..m} {
        let total[i] := total[i] + x[j,i];
    }
}

# Display only non-zero total values
print "     Non-zero clusters";
for {i in 1..m: total[i] > 0} {
    printf "Cluster %d has %d points\n", i, total[i];
}

EOF
```

## 5. A description of how the minimum spanning tree was computed (which software used, etc.)

## 6. The heuristic solution obtained with the minimum spanning tree procedure.

## 7. A comparison of the two solutions obtained, in terms of the objective function of the integer optimization problem. Optionally, you may provide any other additional criteria for comparison.

## 8. Any other observation or comment you may want to add, or any problem you had when performing the assignment.

Conclusion: 
...

# Not to be included:

```{r}
setwd("~/uni_folder/OTDM/UPC-OTDM-main/3/linprog")
#cd /home/julian/uni_folder/OTDM/OTDM_Project_2/3/linprog/
```

(for generating script:
cd /home/julian/uni_folder/OTDM/OTDM_Project_2/3/linprog/
python3 clean_data_linprog.py wine-clustering.csv 3 scaledWine.txt matrixWine.txt ",")





